---
title: "Final Project 274"
author: "Katie Schmitzer"
date: "5/23/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#install.packages("devtools")
#devtools::install_github("FinYang/tsdl")
#install.packages("qpcR")
library(qpcR)
library(tsdl)
#install.packages("tseries")
library(tseries)
library(forecast)
#install.packages("TSA")
require(TSA)
#install.packages("GeneCycle")
```

$\textbf{Get Initial Data}$\newline
```{r}
#tsdl
#for(i in 200:600)  cat("i =", i, attr(tsdl[[i]], "description"), "\n")
chemical <- ts(tsdl[[508]]) #Chemical process readings every two minutes
actualvals<-chemical[91:100]
chem <- chemical[0:90]
chem[30:40]
length(chem)
```

$\textbf{Examine Raw Data}$\newline
```{r}
ts.plot(chem) + abline(h=151.16, col = "red", main = "Chemical Process Readings Every Two Minutes in Degrees Fahrenheit")
mean(chem) # mean is 151.9
var(chem) # variance is 167.6865
acf(chem, main =  "ACF for Raw Data", lag.max = 90) # acf has non-zero values at lags 1 and 2 and ones slightly outside the CI after
pacf(chem, main =  "PACF for Raw Data", lag.max = 90) # pacf has non-zero values at lags 1 and 2 and 20
adf.test(chem)
```

$\textbf{Transform the Data}$\newline
```{r}
bcTransform <- boxcox(chem ~ as.numeric(1:length(chem))) #graph shows that either a log-transformed data or non-transformed data are within 95% CI
lambda <- bcTransform$x[which(bcTransform$y == max(bcTransform$y))]
lambda # ideal lambda = -1.191919
```

$\textbf{Try -1 For Lambda}$\newline
```{r}
lambda = -1 # chose -1 for simplicity and parsimony
chemneg1 <- (1/lambda)*(chem^lambda - 1)
ts.plot(chemneg1, gpars = list(ylab = "Transformed Data")) + abline(h= 0.9933702, col = "red")
mean(chemneg1) # mean is  0.9933702
var(chemneg1) # variance is 3.073924e-07
acf(chemneg1, main =  "ACF for Transformed lambda = -1 Data", lag.max = 40) # acf has non-zero values at lags 1 and 2
pacf(chemneg1, main =  "PACF for Transformed lambda = -1 Data", lag.max = 40) # may be an MA(2) model, non-zero lag at 20


adf.test(chemneg1) # data is stationary
```

$\textbf{Try Log-Data}$\newline
```{r}
chemlog <- log(chem)
ts.plot(chemlog) + abline(h= 5.014343, col = "red")
mean(chemlog) # mean is  5.019678
var(chemlog) # variance is 0.007126066
acf(chemlog, main =  "ACF for Log Data", lag.max =  40) # acf has non-zero values at lags 1 and 2
pacf(chemlog, main =  "PACF for Log Data", lag.max =  40) # may be an MA(2) model
```

$\textbf{De-Trend the Data by differencing at lag 1}$\newline
```{r}
chemneg1_diff1 <-diff(chemneg1,1) 
ts.plot(chemneg1_diff1, main = "Differenced at Lag 1")
mean(chemneg1_diff1) # mean is nearly 0
var(chemneg1_diff1) #  - variance increased
acf(chemneg1_diff1, main =  "ACF for Differenced Data") # acf has more non zero lags
pacf(chemneg1_diff1, main =  "PACF for Differenced Data", lag.max = 40) # pacf has many non zero values
```

$\textbf{Use Innovation Algorithm}$\newline
```{r}
innovations.algorithm <- function(n.max,acvs)
  {
    thetas <- matrix(rep(0,n.max^2),nrow=n.max)
    vs <- rep(0,n.max+1)
    vs[1] <- acvs[1]
    for(n in 1:n.max)
      {
        thetas[n,n] <- acvs[n+1]/vs[1]
        if(n>1)
          {
            for(k in 1:(n-1))
              {
                inner.sum <- acvs[n-k+1]
                for(j in 0:(k-1))
                  {
                    inner.sum <- inner.sum - thetas[k,k-j]*thetas[n,n-j]*vs[j+1]
                  }
                thetas[n,n-k] <- inner.sum/vs[k+1]
              }
          }
        vs[n+1] <- acvs[1]
        for(j in 0:(n-1))
          {
            vs[n+1] <- vs[n+1] - thetas[n,n-j]^2*vs[j+1]
          }
      }
    structure(list(thetas=thetas,vs=vs))
  }

get.sigma.2.n.hs <- function(n,h.max,ts.var,ia.stuff)
  {
    thetas <- ia.stuff$thetas
    vs <- ia.stuff$vs
    sigma.2.n.hs <- rep(ts.var,h.max)
    for(h in 1:h.max)
      {
        sigma.2.n.hs[h] <- sigma.2.n.hs[h] - sum((thetas[n+h-1,(n+h-1):h])^2*vs[1:n])
      }
    sigma.2.n.hs
  }


acvf = acf(chemneg1, plot=FALSE, lag.max = length(chemneg1))$acf[,1,1] * var(chemneg1) 
m = length(acvf)


spec.arma <- function(ar=0,ma=0,var.noise=1,n.freq=500, ...)
{ 
    # check causality
     ar.poly <- c(1, -ar)
     z.ar <- polyroot(ar.poly)
     if(any(abs(z.ar) <= 1)) cat("WARNING: Model Not Causal", "\n")  
    # check invertibility
     ma.poly <- c(1, ma)
     z.ma <- polyroot(ma.poly)
     if(any(abs(z.ma) <= 1)) cat("WARNING: Model Not Invertible", "\n")
     if(any(abs(z.ma) <= 1) || any(abs(z.ar) <= 1) ) stop("Try Again")
    #
    ar.order <- length(ar)
    ma.order <- length(ma) 
    # check (near) parameter redundancy [i.e. are any roots (approximately) equal]  
       for (i in 1:ar.order) {
       if ( (ar == 0 & ar.order == 1) || (ma == 0 & ma.order ==1) ) break
       if(any(abs(z.ar[i]-z.ma[1:ma.order]) < 1e-03)) {cat("WARNING: Parameter Redundancy", "\n"); break}
       }
    #
    freq <- seq.int(0, 0.5, length.out = n.freq)
            cs.ar <- outer(freq, 1:ar.order, function(x, y) cos(2 * 
                pi * x * y)) %*% ar
            sn.ar <- outer(freq, 1:ar.order, function(x, y) sin(2 * 
                pi * x * y)) %*% ar
            cs.ma <- outer(freq, 1:ma.order, function(x, y) cos(2 * 
                pi * x * y)) %*% -ma
            sn.ma <- outer(freq, 1:ma.order, function(x, y) sin(2 * 
                pi * x * y)) %*% -ma                      
    spec <- var.noise*((1 - cs.ma)^2 + sn.ma^2)/((1 - cs.ar)^2 + sn.ar^2)
    spg.out <- list(freq=freq, spec=spec)
    class(spg.out) <- "spec"
    plot(spg.out, ci=0, main="", ...)
    return(invisible(spg.out))
}
#chemneg1.ia = innovations.algorithm(m+1, acvf)
#chemneg1.ia$thetas[3,1:3] 
```

$\textbf{Fit to an AR Model}$\newline
```{r}
fit.ar.yw <- ar(chemneg1, aic = TRUE, order.max = NULL, method = c("yule-walker")) 
fit.ar.yw # suggests AR(2) with the coefficients : 0.4085  -0.4117  
0.4085+1.96*2.393e-07
0.4085-1.96*2.393e-07
-0.4117+1.96*2.393e-07
-0.4117-1.96*2.393e-07
fit.ar.mle <- ar(chemneg1, aic = TRUE, order.max = NULL, method = c("mle")) 
fit.ar.mle  # suggests AR(2) with the coefficients : 0.4041  -0.4211  

```

$\textbf{Fitting to ARMA Model}$\newline
```{r}
aiccs <- matrix(NA, nr = 9, nc = 9)
dimnames(aiccs) = list(p=0:8, q=0:8)
for(p in 0:8)
{
  for(q in 0:8)
  {
aiccs[p+1,q+1] = AICc(arima(chemneg1, order = c(p,0,q), method="ML")) }
}

aiccs # possible models include ARMA(1,3), ARMA(0,2), ARMA(0,3), ARMA(2,1), ARMA(3,5), ARMA(1,3), ARMA(2,0) even though it has the lowest AICc because it is much too high an order. The ACF and PACF plot suggest an AR(2) model.

min(aiccs) # ARMA(1,3)

auto.arima(chemneg1, stepwise = F, approximation = F, lambda = -1, ic = "aicc")
```

$\textbf{Trying ARMA(2,0)}$\newline
```{r}
arma20 <- arima(chemneg1, order = c(2, 0, 0), method = c("ML"))
arma20
fit.ar.mle

mean(residuals(arma20)) 
plot(residuals(arma20), main = "Residuals of ARMA(2,0) Model")
Box.test(residuals(arma20), lag = sqrt(90), type = "Ljung-Box", fitdf = 2) 
Box.test(residuals(arma20), type = "Box-Pierce", lag = sqrt(90), fitdf = 2) 
Box.test(residuals(arma20)^2, type = "Ljung", lag = sqrt(90), fitdf = 0) 

acf(residuals(arma20),main = "Autocorrelation")
pacf(residuals(arma20),main = "Partial Autocorrelation")
residuals(arma20)
hist(residuals(arma20),main = "Histogram") 
qqnorm(residuals(arma20))  
qqline(residuals(arma20),col ="blue")
shapiro.test(residuals(arma20))
fit.ar <- ar(residuals(arma20),method="yw")
fit.ar
tsdiag(arima(chemneg1, order=c(2,0,0))) 
```

$\textbf{Trying ARMA(1,3)}$\newline
```{r}
arma13 <- arima(chemneg1, order = c(1, 0, 3), method = c("ML"))
arma13
-0.6922+1.96*0.1286 #ar1         
-0.6922-1.96*0.1286
1.2634+1.96*0.1393 #ma1 
1.2634-1.96*0.1393
0.0446+1.96*0.1889 #ma2 contains 0
0.0446-1.96*0.1889 
-0.4117+1.96*0.1039 #ma3
-0.4117-1.96*0.1039 

spec.arma(ar=c( -0.6946),ma=c(1.2634,0.0446,-0.4117)) 

plot(residuals(arma13), main = "Residuals of ARMA(1,3) Model") + abline(h = mean(residuals(arma13)), col = "red")
mean(residuals(arma13)) 
Box.test(residuals(arma13), type = "Ljung-Box", lag = sqrt(90), fitdf = 4) 
Box.test(residuals(arma13), type = "Box-Pierce", lag = sqrt(90), fitdf = 4) 
Box.test((residuals(arma13))^2, type = "Ljung-Box", lag = sqrt(90), fitdf = 0) 

acf(residuals(arma13),main = "Autocorrelation", lag.max = 40) 
pacf(residuals(arma13),main = "Partial Autocorrelation", lag.max = 40)
residuals(arma13)
hist(residuals(arma13),main = "Histogram") 
qqnorm(residuals(arma13)) 
qqline(residuals(arma13),col ="blue")
shapiro.test(residuals(arma13))
tsdiag(arima(chemneg1, order=c(1,0,3))) 
```

$\textbf{Trying ARMA(2,1)}$\newline
```{r}
arma21 <- arima(chemneg1, order = c(2, 0, 1), method = c("ML"))
arma21
-0.0077+1.96*0.1879 #ar1 contains 0
-0.0077-1.96*0.1879
-0.2791+1.96*0.1316
-0.2791-1.96*0.1316
0.5145+1.96*0.1875
0.5145-1.96*0.1875

spec.arma(ar=c( -0.0077,-0.2791),ma=c(0.5145))

plot(residuals(arma21), main = "Residuals of ARMA(2,1) Model")
mean(residuals(arma21))
Box.test(residuals(arma21), type = "Ljung-Box", lag = sqrt(90), fitdf = 3) 
Box.test(residuals(arma21), type = "Box-Pierce", lag = sqrt(90), fitdf = 3)  
Box.test(residuals(arma21)^2, type = "Ljung-Box", lag = sqrt(90), fitdf = 0) 

acf(residuals(arma21),main = "Autocorrelation")
pacf(residuals(arma21),main = "Partial Autocorrelation")
residuals(arma21)
hist(residuals(arma21),main = "Histogram") 
qqnorm(residuals(arma21)) 
qqline(residuals(arma21),col ="blue")
shapiro.test(residuals(arma21))
tsdiag(arima(chemneg1, order=c(2,0,1))) 
```

$\textbf{Trying ARMA(0,3)}$\newline
```{r}
arma03 <- arima(chemneg1, order = c(0, 0, 3), method = c("ML"))
arma03
0.5137+1.96*0.1086
0.5137-1.96*0.1086
-0.2883+1.96*0.1048
-0.2883-1.96*0.1048
-0.1797+1.96*0.1007 # ma3 contains 0
-0.1797-1.96*0.1007


spec.arma(ar=c(0),ma=c(0.5137,-0.2883, -0.1797)) 

mean(residuals(arma03)) 
plot(residuals(arma03), main = "Residuals of ARMA(0,3) Model")
Box.test(residuals(arma03), lag = sqrt(90), type = "Ljung-Box", fitdf = 3) 
Box.test(residuals(arma03), type = "Box-Pierce", lag = sqrt(90), fitdf = 3) 
Box.test(residuals(arma03)^2, type = "Ljung", lag = sqrt(90), fitdf = 0) 


acf(residuals(arma03),main = "Autocorrelation")
pacf(residuals(arma03),main = "Partial Autocorrelation")
residuals(arma03)
hist(residuals(arma03),main = "Histogram") 
qqnorm(residuals(arma03)) 
qqline(residuals(arma03),col ="blue")
shapiro.test(residuals(arma03)) 
fit.ar <- ar(residuals(arma03),method="yw")
fit.ar
tsdiag(arima(chemneg1, order=c(0,0,3))) 
```

$\textbf{Trying ARMA(0,2)}$\newline
```{r}
arma02 <- arima(chemneg1, order = c(0, 0, 2), method = c("ML"))
arma02
0.4548+1.96*0.1219
0.4548-1.96*0.1219
-0.2093+1.96*0.1196 
-0.2093-1.96*0.1196
mean(residuals(arma02))
plot(residuals(arma02), main = "Residuals of ARMA(0,2) Model")
Box.test(residuals(arma02), type = "Ljung-Box", lag = sqrt(90), fitdf = 2) 
Box.test(residuals(arma02), type = "Box-Pierce", lag = sqrt(90), fitdf = 2) 
Box.test(residuals(arma02)^2, type = "Ljung", lag = sqrt(90), fitdf = 0) 

acf(residuals(arma02),main = "Autocorrelation", lag.max = 40)
pacf(residuals(arma02),main = "Partial Autocorrelation", lag.max = 40)
residuals(arma02)
hist(residuals(arma02),main = "Histogram") 
qqnorm(residuals(arma02)) 
qqline(residuals(arma02),col ="blue")
shapiro.test(residuals(arma02)) 
tsdiag(arima(chemneg1, order=c(0,0,2))) 
```

$\textbf{Trying ARMA(0,1)}$\newline
```{r}
arma01 <- arima(chemneg1, order = c(0, 0, 1), method = c("ML"))
arma01
0.6349+1.96*0.0821
0.6349-1.96*0.0821
spec.arma(ma=c(0.6349))

mean(residuals(arma01)) 
plot(residuals(arma01), main = "Residuals of ARMA(0,2) Model")
Box.test(residuals(arma01), type = "Ljung-Box", lag = sqrt(90), fitdf = 2) 
Box.test(residuals(arma01), type = "Box-Pierce", lag = sqrt(90), fitdf = 2) 
Box.test(residuals(arma01)^2, type = "Ljung", lag = sqrt(90), fitdf = 0) 

acf(residuals(arma01),main = "Autocorrelation", lag.max = 40)
pacf(residuals(arma01),main = "Partial Autocorrelation", lag.max = 40)
residuals(arma01)
hist(residuals(arma01),main = "Histogram") 
qqnorm(residuals(arma01)) 
qqline(residuals(arma01),col ="blue")
shapiro.test(residuals(arma01))
tsdiag(arima(chemneg1, order=c(0,0,2))) 
```

$\textbf{Predicting 10 Steps Ahead using ARMA(1,3) on transformed data}$\newline
```{r}
pred = predict(arma13, n.ahead=10) 
pred$pred
ts.plot(chemneg1,xlim=c(0,100)) 
points(91:100,pred$pred, col = "blue") 
lines(91:100,pred$pred-1.96*pred$se,lty=2, col = "orange") 
lines(91:100,pred$pred+1.96*pred$se,lty=2, col = "orange")
```

$\textbf{Predicting 10 Steps Ahead using ARMA(1,3) on raw data}$\newline
```{r}
pred = predict(arma13, n.ahead=10) 
pred_raw <- -1/(pred$pred-1)
pred_raw
utran <- pred$pred+1.96*pred$se
ltran <- pred$pred-1.96*pred$se
u <- -1/(utran-1)
l <- -1/(ltran-1)
ts.plot(chem,xlim=c(0,100)) 
points(91:100,pred_raw, col = "red") 
lines(91:100,u,lty=2, col = "orange") 
lines(91:100,l,lty=2, col = "orange")

ts.plot(chem,xlim=c(0,100)) 
points(91:100,pred_raw, col = "red") 
points(91:100,actualvals, col = "black")
lines(91:100,u,lty=2, col = "orange") 
lines(91:100,l,lty=2, col = "orange")

ts.plot(chemical)
```

$\textbf{Spectral Analysis}$\newline
```{r}
periods = periodogram(chemneg1); abline(h=0)
periods

cpgram(residuals(arma13))
FREQ = periods$freq[order(periods$spec, decreasing=T)][1] 
FREQ[1]
```

```{r}
library(GeneCycle)
fisher.g.test(residuals(arma13))
```